---
title: "Practice Dates"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#rmarkdown::run("practice-dates.Rmd")
library(learnr)
knitr::opts_chunk$set(echo= FALSE)
```

# Practice Dates

In R, there are special `Date` objects that help you work with dates and times! `Date` objects have special powers that makes complicated calculations easier. For instance, you can calculate how many days there are between events.

**To practice, let's calculate how many days old you are!**

Before we can accomplish that, we need to understand the toolkit of functions we have available to work with dates.

### Toolkit: `as.Date()`

The `as.Date(...)` function takes a character string in a year-month-day format and turns it into a `Date` object.

Press the blue "run" button to see the output of this code:

```{r as_date, exercise = TRUE}
andrey_birthday <- as.Date("1999-04-20")
print(andrey_birthday)
```

Although the result appears pretty boring, this is not just any character string... it is really a `Date` object!

Try changing the date to your own birthdate, and rename the variable name appropriately.

### Toolkit: `Sys.Date()`

The `Sys.Date()` function returns a `Date` object for today's date.

```{r sys_date, exercise = TRUE, exercise.eval = TRUE}
today_date <- Sys.Date()
print(today_date)
```

Perfect, `today_date` is a `Date` object containing today's date!

### Toolkit: Subtracting Dates

We can subtract two `Date` objects to figure out how many days are in between them.

```{r subtract_dates, exercise = TRUE, exercise.eval = TRUE}
end_of_spring_finals <- as.Date("2019-06-14")
start_of_spring_qtr <- as.Date("2019-04-01")

# How many days in are in spring quarter?
print(end_of_spring_finals - start_of_spring_qtr)
```

Wow, R figures out how many days are in between our two `Date` objects, and even provides nice contextual text when printing out the answer! _(You don't need to know this, but the contextual text shows up because subtracting two `Date` objects creates a `difftime` object. When you print a `difftime` object, it is always formatted to say "Time difference of {X}". Convenient!)._

When subtracting two dates, it can be confusing to figure out which date to subtract from which. One way to remember: you want a _positive_ number for how many days are between the dates, so you want to subtract the _smaller_ number (earlier date) from the _bigger_ number (later date).

### Toolkit: Formatting Dates

If you have a `Date` object and you want to display it in a human-readable way, you can use the `format(...)` function and specify the format you want it in.

For example:

```{r format_date_example, exercise = TRUE, exercise.eval = TRUE}
andrey_birthday <- as.Date("1999-04-20")
formatted_andrey_birthday <- format(andrey_birthday, format = "%B %d, %Y")
print(paste("Andrey's birthdate is", formatted_andrey_birthday))
```

You can format the date in many different ways! Here is a table of all the symbols you can use to format your dates:

| Symbol 	| Meaning                	| Example 	|
|--------	|------------------------	|---------	|
| %d     	| day as a number (0-31) 	| 01      	|
| %a     	| abbreviated weekday    	| Mon     	|
| %A     	| unabbreviated weekday  	| Monday  	|
| %m     	| month (01-12)          	| 01      	|
| %b     	| abbreviated month      	| Jan     	|
| %B     	| unabbreviated month    	| January 	|
| %y     	| 2-digit year           	| 19      	|
| %Y     	| 4-digit year           	| 2019    	|

To practice, format today's date in a unique way:

```{r format_date_exercise, exercise = TRUE}
today_date <- Sys.Date()
# print()
```

### Exercise: How Many Days Old Am I?

Let's do an exercise in calculating how many days old you are to practice working with `Date` objects! Here's our prompt:

> Our objective is to write a function `get_age_days` that takes in one argument `my_birthday` that is a `Date` object. The function should return a character string in the format `"I am {X} days old". Finally, test your function by providing your own birthdate.

Complete the exercise here:

```{r get_age_days_ex, exercise = TRUE}
# Create a function `get_age_days` that takes in one argument `my_birthday`
# and returns a character string in the format `"I am {X} days old"
# Hint: you will need to use the `paste(...)` function to combine strings.


# Create a variable representing your birthdate


# Use the `print` function to output the result

```

# Congratulations!

Wonderful! You now have practice in:

- Manipulating dates
- Writing functions
- Combining strings

Amazing!! You're going to do great on A2 and the whole rest of the quarter :-)

